---
title: "Monte Carlo Simulation: Part 1"
output: html_document
---

```{r setup, include = FALSE}
# packages required for this post
for (pkg in c('tidyquant', 'tidyverse', 'timetk', 'tibbletime', 'broom')) 
  if (!requireNamespace(pkg)) install.packages(pkg)

knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment= NA)
```

Today we start a series of posts on running and visualizing Monte Carlo portfolio simulations. This series is excerpted from my new book [Reproducible Finance with R](https://www.amazon.com/Reproducible-Finance-Portfolio-Analysis-Chapman/dp/1138484032)

Let's get to it!

Monte Carlo relies on repeated, random sampling and we will sample based on two parameters: mean and standard deviation of portfolio returns.

    + SPY (S&P500 fund) weighted 25%
    + EFA (a non-US equities fund) weighted 25%
    + IJS (a small-cap value fund) weighted 20%
    + EEM (an emerging-mkts fund) weighted 20%
    + AGG (a bond fund) weighted 10%

Before we can simulate that portfolio, we need to calculate the historical portfolio monthly returns, which was covered in this article on [Introduction to Portfolio Returns](http://www.reproduciblefinance.com/2017/10/12/introduction-to-portfolio-returns/). 

I won't go through the logic again, but the code is here: 

```{r}
# These are the package we need for today's post.
library(tidyquant)
library(tidyverse)
library(timetk)
library(broom)

symbols <- c("SPY","EFA", "IJS", "EEM","AGG")

prices <- 
  getSymbols(symbols, src = 'yahoo', 
             from = "2012-12-31",
             to = "2017-12-31",
             auto.assign = TRUE, warnings = FALSE) %>% 
  map(~Ad(get(.))) %>%
  reduce(merge) %>% 
  `colnames<-`(symbols)

w <- c(0.25, 0.25, 0.20, 0.20, 0.10)

asset_returns_long <-  
  prices %>% 
  to.monthly(indexAt = "lastof", OHLC = FALSE) %>% 
  tk_tbl(preserve_index = TRUE, rename_index = "date") %>%
  gather(asset, returns, -date) %>% 
  group_by(asset) %>%  
  mutate(returns = (log(returns) - log(lag(returns)))) %>% 
  na.omit()

portfolio_returns_tq_rebalanced_monthly <- 
  asset_returns_long %>%
  tq_portfolio(assets_col  = asset, 
               returns_col = returns,
               weights     = w,
               col_rename  = "returns",
               rebalance_on = "months")
```

We will be working with the data object `portfolio_returns_tq_rebalanced_monthly` and we first find the mean and [standard deviation](http://www.reproduciblefinance.com/code/standard-deviation/) of returns.

We will name those variables  `mean_port_return` and `stddev_port_return`. 

```{r}
mean_port_return <- 
  mean(portfolio_returns_tq_rebalanced_monthly$returns)

stddev_port_return <- 
  sd(portfolio_returns_tq_rebalanced_monthly$returns)
```


Then we use the `rnorm()` function to sample from a distribution with mean equal to `mean_port_return` and standard deviation equal to `stddev_port_return`.  That is the crucial random sampling that underpins this exercise. 

We also must decide how many draws to pull from this distribution, meaning how many monthly returns we will simulate.  120 months is 10 years and that feels like a good amount of time.

```{r}
simulated_monthly_returns <- rnorm(120, 
                                   mean_port_return, 
                                   stddev_port_return)
```

Have a quick look at the simulated monthly returns. 

```{r}
head(simulated_monthly_returns)
tail(simulated_monthly_returns)
```

Next, we calculate how a dollar would have grown given those random monthly returns. We first add a 1 to each of our monthly returns, because we start with $1.

```{r}
simulated_returns_add_1 <- 
  tibble(c(1, 1 + simulated_monthly_returns)) %>% 
  `colnames<-`("returns")

head(simulated_returns_add_1)
```

That data is now ready to be converted into the cumulative growth of a dollar. We can use either `accumulate()` from `purrr` or `cumprod()`. Let's use both of them with `mutate()` and confirm consistent, reasonable results. 

```{r}

simulated_growth <- 
simulated_returns_add_1 %>%
    mutate(growth1 = accumulate(returns, function(x, y) x * y),
           growth2 = accumulate(returns, `*`),
           growth3 = cumprod(returns)) %>% 
    select(-returns)

tail(simulated_growth)
```

We just ran 3 simulations of dollar growth over 120 months. We passed in the same monthly returns and that's why we got 3 equivalent results.

Are they reasonable? What compound annual growth rate (CAGR) is implied by this simulation?

```{r}
cagr <- 
  ((simulated_growth$growth1[nrow(simulated_growth)]^
      (1/10)) - 1) * 100

cagr <- round(cagr, 2)
```


This simulation implies an annual compounded growth of `r cagr`%. That seems reasonable given our actual returns have all been taken from a raging bull market. Remember, the above code is a simulation based on sampling from a normal distribution. If you re-run this code on your own, you will get a different result.

If we feel good about this first simulation, we can run several more to get a sense for how they are distributed. Before we do that, let's create several different functions that could run the same simulation.

## Several Simulation Functions 

Let's build 3 simulation functions that incorporate the `accumulate()` and `cumprod()` workflows above. We have confirmed they give consistent results so it's a matter of stylistic preference as to which one is chosen in the end. Perhaps you feel that one is more flexible or extensible or fits better with your team's code flows.

Each of the below functions needs 4 arguments: N for the number of months to simulate (we chose 120 above), `init_value` for the starting value (we used $1 above) and the mean-standard deviation pair to create draws from a normal distribution. We *choose* N and `init_value`, and derive the mean-standard deviation pair from our portfolio monthly returns. 

Here is our first growth simulation function using `accumulate()`.

```{r}

simulation_accum_1 <- function(init_value, N, mean, stdev) {
    tibble(c(init_value, 1 + rnorm(N, mean, stdev))) %>% 
    `colnames<-`("returns") %>%
    mutate(growth = 
             accumulate(returns, 
                        function(x, y) x * y)) %>% 
    select(growth)
}
```

Almost identical, here is the second simulation function using `accumulate()`.

```{r}

simulation_accum_2 <- function(init_value, N, mean, stdev) {
  tibble(c(init_value, 1 + rnorm(N, mean, stdev))) %>% 
    `colnames<-`("returns") %>%
  mutate(growth = accumulate(returns, `*`)) %>% 
  select(growth)
}
```

Finally, here is a simulation function using `cumprod()`.

```{r}
simulation_cumprod <- function(init_value, N, mean, stdev) {
  tibble(c(init_value, 1 + rnorm(N, mean, stdev))) %>% 
    `colnames<-`("returns") %>%
  mutate(growth = cumprod(returns)) %>% 
  select(growth)
}
```

Here is a function that uses all three methods, in case we want a fast way to re-confirm consistency. 

```{r}

simulation_confirm_all <- function(init_value, N, mean, stdev) {
  tibble(c(init_value, 1 + rnorm(N, mean, stdev))) %>% 
    `colnames<-`("returns") %>%
    mutate(growth1 = accumulate(returns, function(x, y) x * y),
           growth2 = accumulate(returns, `*`),
           growth3 = cumprod(returns)) %>% 
    select(-returns)
}
```

Let's test that `confirm_all()` function with an `init_value` of 1, N of 120, and our parameters.

```{r}
simulation_confirm_all_test <- 
  simulation_confirm_all(1, 120, 
                         mean_port_return, stddev_port_return)

tail(simulation_confirm_all_test)
```

Consistent results and we're ready to visualize! 

That's all for today. Next time we will explore methods for running more than one simulation with the above functions and then charting the results.


And....

If you've read this far, good news! Between now and December 31, 2018, there's a 20% discount on the book being run at [CRC](https://crcpress.com/Reproducible-Finance-with-R-Code-Flows-and-Shiny-Apps-for-Portfolio-Analysis/Jr/p/book/9781138484030), and if you don't see it applied, readers can use discount code SS120 on the [CRC website](https://crcpress.com/Reproducible-Finance-with-R-Code-Flows-and-Shiny-Apps-for-Portfolio-Analysis/Jr/p/book/9781138484030). The book is also available on [Amazon as Kindle or paperback](https://www.amazon.com/Reproducible-Finance-Portfolio-Analysis-Chapman/dp/1138484032).

Thanks for reading!